---
title: "tutorial stuff"
output:
  html_document: default
  html_notebook: default
---

# 1 DT

An interactive data chart.

```{r}
if(!require(DT)) install.packages("DT")
library(DT)
datatable(iris) # interactive data chart
```

# 2 data.table::fread

Fast reading.

```{r}
if(!require(data.table)) install.packages("data.table")
library(data.table)

?fread # omg fread is so freaking fast
```

# 3 Mutilple ways to present data on a US map

## 3.1 chorolepth

* pros:
    1. uses straightfoward index (lower case state name and 
    2. with legend5 digit fips code)
    3. possible zooming into specific part of map
* cons:
    1. static
    2. only US state map and US county map


```{r}

library(choroplethr)
library(choroplethrMaps)

?df_pop_state
data(df_pop_state)

?state_choropleth # static maps
state_choropleth(df_pop_state) # uses lower case state name as index

data(df_pop_county)
county_choropleth(df_pop_county,
                  title       = "Population",
                  legend      = "Population",
                  num_colors  = 1,
                  state_zoom  = "new york") # uses 5 digit fips code as index

```


## 3.2 plotly

* pros
    1. good looking
    2. interactive with hovers, zoom in and out
* cons:
    1. no documentation, how to adjust parameters?

```{r}
if(!require(plotly)) install.packages("plotly")
library(plotly)

# plotly state maps https://plot.ly/r/choropleth-maps/ interactive maps

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

plot_ly(df, z = total.exports, text = hover, locations = code, type = 'choropleth',
        locationmode = 'USA-states', color = total.exports, colors = 'Purples',
        marker = list(line = l), colorbar = list(title = "Millions USD")) %>%
  layout(title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)', geo = g)

```

## 3.3 leaflet + maps

See http://rstudio.github.io/leaflet/

* pros
    1. can apply to anywhere in the world
    2. interactive with detailed map, zoom in and out
* cons:
    1. rely on predefined state or county maps in "maps" package
    2. there are other ways to add state/county polygons but not straight forward


```{r}
if(!require(maps)) install.packages("maps")
if(!require(leaflet)) install.packages("leaflet")

library(maps)
library(leaflet)

mapStates <- map(database = "state", fill = TRUE, plot = FALSE)
pal <- colorNumeric(palette = "OrRd", domain = 1)
leaflet(data = mapStates) %>% addTiles() %>%
  addPolygons(fillColor = pal(1), fillOpacity=0.5,stroke = F)

```

```{r}
library(leaflet)
m = leaflet() %>% addTiles()  # a map with the default OSM tile layer
m
m %>% fitBounds(0, 40, 10, 60) # set bounds Europe map
m %>% setView(-73.9619572,40.8078214, zoom = 17) # move the center to Columbia university

```

